---
title: Matrix Approximation with Kronecker and Precision Weighting
author: 
    - name: Jan de Leeuw
      orcid: 0000-0003-1420-1797
      email: jan@deleeuwpdx.net
      affiliation: 
        - name: University of California Los Angeles
          city: Los Angeles
          state: CA
          url: www.ucla.edu
    - name: Jan Graffelman
license: "CC0"
date: last-modified
date-format: long
bibliography: [mypubs.bib, total.bib]
number-sections: true
pdf-engine: lualatex
keep-tex: true
format:
   pdf:
    fontsize: 12pt
    include-in-header: 
     - preamble.tex
    keep-tex: true
    link-citations: true
    documentclass: scrartcl
    number-sections: true
   html:
    fontsize: 12pt
    include-in-header: 
     - preamble.css
    keep-md: true
    number-sections: true
toc: true
toc-depth: 3
editor: source
papersize: letter
graphics: true
link-citations: true
mainfont: Times New Roman
abstract: TBD
---

```{r load_packages, echo = FALSE}
suppressPackageStartupMessages(library(knitr, quietly = TRUE))
suppressPackageStartupMessages(library(tinytex, quietly = TRUE))
suppressPackageStartupMessages(library(microbenchmark, quietly = TRUE))
```

```{r load_sources, echo = FALSE}
source("triple.R")
source("nested.R")
source("checker.R")
```

\sectionbreak 

**Note:** This is a working manuscript which will be expanded/updated
frequently. All suggestions for improvement are welcome. All Rmd, tex,
html, pdf, R, and C files are in the public domain. Attribution will be
appreciated, but is not required. The files can be found at
<https://github.com/deleeuw/triple> 

\sectionbreak

# Introduction

The problem addressed in this note is to approximate a given $n\times m$ matrix $Y$
by a $n\times m$ matrix $X$ which must be in a given subset $\mathbb{X}$ of
$\mathbb{R}^{n\times m}$. For example, $\mathbb{X}$ can be the $n\times m$
matrices of rank less than or equal to $p$, or a set of special matrices, such as
the Hankel or Cauchy matrices.

We measure loss using a loss function proposed in @graffelman_26a.
\begin{equation}
\sigma(X):=\text{tr}\ R(W\circ(Y-X))C(W\circ(Y-X))'.\label{eq-loss}
\end{equation}
In Definition \eqref{eq-loss} $R$ and $C$ are given positive semi-definite matrices of orders $n$ and $m$, which define the *Kronecker weights*. The Kronecker weights can be, for example, inverse covariance matrices for row and column entries, as in the matrix variate normal distribution (@gupta_nagar_00). The $n\times m$ matrix $W$ of *precision weights* has positive entries, which can be, for example, inverse standard errors of the corresponding entries of $Y$. The symbol $\circ$ is used for Hadamard (elementwise) multiplication.

Loss function \eqref{eq-loss} has many well-known special cases. If all weights
are equal to one then we recover correspondence analysis, redundancy analysis,
and various other forms of canonical analysis, as well as 
forms of principal component analysis with Kronecker weights (@deleeuw_A_84b, @deleeuw_kukuyeva_U_09, @graffelman_26a, @graffelman_26b). 
If both $R$ and $C$ are equal to the identity we have principal components analysis with precision weights, as in @gabriel_zamir_79. If $Y$ is symmetric the various approximations of the correlation matrix discussed in @graffelman_deleeuw_A_23 are special cases.


\sectionbreak

# Algorithms

Our algorithms use majorization. See @lange_16 or @deleeuw_B_21c for comprehensive discussions of majorization (or MM).
Majorization is used to reduce the weighted least squares problem to an unweighted
least squares problem, which is generally much easier to solve. Similar applications 
of MM are in @kiers_97 and in @groenen_giaquinto_kiers_03. 

## Simple Majorization

We first *vectorize* the problem. This means replacing the definition \eqref{eq-loss}
that uses matrices by an equivalent definition in terms of vectors, which is easier to work with. Define
$x:=\text{vec}(X)$ and $y:=\text{vec}(X)$. Loss becomes
\begin{equation}
\sigma(x):=(y-x)'W_D(C\otimes R)W_D(y-x).\label{eq-lossv}
\end{equation}
In \eqref{eq-lossv} matrix $W_D$ is the diagonal matrix with the
elements of $\text{vec}(W)$ on the diagonal.

Suppose $\lambda$ is any number larger than or equal to the largest
eigenvalue $\lambda_{\text{max}}(H)$ of $H:=W_D(C\otimes R)W_D$. Then, for all $x$ and $\tilde x$ substitute $x=\tilde x+(x-\tilde x)$ in \eqref{eq-lossv}. This gives
\begin{align}
\sigma(x)&=((y-\tilde x)-(x-\tilde x))'H((y-\tilde x)-(x-\tilde x))\notag\\
&=\sigma(\tilde x)-2(x-\tilde x)'H(y-\tilde x)+(x-\tilde x)'H(x-\tilde x)\notag\\
&\leq\sigma(\tilde x)-2(x-\tilde x)'H(y-\tilde x)+\lambda\|x-\tilde x\|^2\notag\\
&=\sigma(\tilde x)+\lambda\|x-\tilde y\|^2-\lambda^{-1}\|H(y-\tilde x)\|^2,
\end{align}
with 
\begin{equation}
\tilde y:=\tilde x+\lambda^{-1}H(y-\tilde x).
\end{equation}

A majorization (MM) step minimizes $\|x-\tilde y\|^2$ over the $x$ satisfying the
constraints. If we now *matricize* (the inverse of vectorize) it follows that we have to minimize $\|X-\tilde Y\|^2$,
over $\mathbb{X}$ with
\begin{equation}
\tilde Y:=\tilde X+\lambda^{-1}W\circ\{R(W\circ(Y-\tilde X))C\}.
\end{equation}
Note that this expression for $\tilde Y$ does not use the (possibly very large)
Kronecker product $H$. If computing the largest eigenvalue $\lambda_{\text{max}}(H)$ (which only has to be done once before the iterations start) is a problem then we can use an upper bound.
\begin{multline}
\max_z\frac{z'W_D(C\otimes R)W_Dz}{z'z}=\max_z\frac{z'W_D(C\otimes R)W_Dz}{z'W_D^2z}\frac{z'W_D^2z}{z'z}\\
\leq\max_z\frac{z'W_D(C\otimes R)W_Dz}{z'W_D^2z}\max_z\frac{z'W_D^2z}{z'z}
=\lambda_{\text{max}}(C)\lambda_{\text{max}}(R)\max(W\circ W)
\end{multline}

## Nested Majorization

There is a more radical way to avoid computing the largest eigenvalue of
$H$. We employ two nested majorizations. In the first majorization
we get rid of the Kronecker weighting and reduce each MM iteration to a
weighted least squares problem that is separable, i.e. has a diagonal weight matrix. 
The second majorization is used to make
a further reduction of this separable weighted least squares problem
to a sequence of unweighted least squares problems. The majorizations are nested, 
in the sense that we need one or more iterations of the second majorization process
within each iteration of the first majorization process.

### First Majorization

Start with the same vectorization we used before in our simple majorization But instead of
majorizing by using $z'Hz\leq\lambda_{\text{max}}(H)z'z$ we use
\begin{equation}
z'Hz\leq\lambda_{\text{max}}(C\otimes R)z'W_D^2z=\lambda_{\text{max}}(C)\lambda_{\text{max}}(R)z'W_D^2z.
\end{equation}
Let $\lambda:=\lambda_{\text{max}}(C)\lambda_{\text{max}}(R)$. We now have
\begin{align}
\sigma(x)&=((y-\tilde x)-(x-\tilde x))'H((y-\tilde x)-(x-\tilde x))\notag\\
&=\sigma(\tilde x)-2(x-\tilde x)'H(y-\tilde x)+(x-\tilde x)'H(x-\tilde x)\notag\\
&\leq\sigma(\tilde x)-2(x-\tilde x)'H(y-\tilde x)+\lambda(x-\tilde x)'W_D^2(x-\tilde x)\notag\\
&=\sigma(\tilde x)+\lambda(x-\tilde y)'W_D^2(x-\tilde y)+\lambda^{-1}(y-\tilde x)'HW_D^{-2}H(y-\tilde x)
\end{align}

\begin{equation}
\tilde y:=\tilde x+\lambda^{-1}W_D^{-2}H(y-\tilde x),
\end{equation}
which can be matricized to
\begin{equation}
\tilde Y:=\tilde X+\lambda^{-1}W^{-1}\circ\{R(W\circ(Y-\tilde X))C\}.
\end{equation}

### Second majorization

In the "inner" majorization we need to minimize
\begin{equation}
\sum_{i=1}^n\sum_{j=1}^nw_{ij}^2(\tilde y_{ij}-x_{ij})^2\label{eq-inloss}
\end{equation}
over $X\in\mathbb{X}$. Note that in the inner iterations the "target"
$\tilde Y$ is the same in each iteration, because it only changes in the 
outer iterations. Minimizing (or decreasing) \eqref{eq-inloss} can be done with the wAddPCA() function from the Correlplot package (@graffelman_deleeuw_M_26).

The majorization for the inner iterations is simple. The same quadratic majorization reasoning as in the previous sections leads to the unweighted problem of minimizing
\begin{equation}
\sum_{i=1}^n\sum_{j=1}^n(\tilde{\tilde y}_{ij}-x_{ij})^2,
\end{equation}
with
\begin{equation}
\tilde{\tilde y}_{ij}:=\tilde x_{ij}+\frac{w_{ij}^2}{w^2_{\text{max}}}(\tilde y_{ij}-\tilde x_{ij}).
\end{equation}

\sectionbreak

# Software

The github repository at <https://github.com/deleeuw/triple> has the file triple.R, which constains the function triple(), for simple majorization, and the file nested.R with the R program nested(), for nested majorization. 

## Arguments 

Arguments are
```{r args_triple}
args(triple)
```
```{r args_nested}
args(nested)
```
The arguments are mostly self-explanotory. An exception is perhaps the func parameter.
which is a separate R routine that performs the unweighted least squares projection on $\mathbb{X}$. Parameter $p$ has additional parameters for func, if needed. The default
uses the truncated SVD for low-rank matrix approximation with $p=2$. The initial estimate is the unweighted projection of the data matrix, without taking any of the weight matrices into account. Note that nested() has stopping parameters itmax and eps for the outer iterations and inmax and ips for the inner iterations.

## Values

## Projection Routines

The file auxiliary.R has routines for unweighted least squares
projection in matrix space. This defines possible values for the
func parameter of triple() and nested().

* eckart_young() low-rank approximation.
* column_adjust() subtract column means and then eckart_young.
* row_adjust() subtract row means and then eckart_young
* double_center() and then eckart_young.
* hankel() finds the best fitting Hankel matrix (square matrices only).
* toeplitz() finds the best fitting Toeplitz matrix (square matrices only).
* symmetric() finds the best fitting symmetric matrix (square matrices only).
* anti_symmetric() finds the best fitting anti-symmetric matrix (square matrices only).
* finds the best fitting additive matrix (i.e. with $x_{ij}=a_i+b_j$)


\sectionbreak

# Example

We analyze a small artifical example, generated from random data.

## Data

```{r}
ymat
```
```{r}
rmat
```
```{r}
cmat
```
```{r}
wmat
```

## Triple

```{r}
h <- triple(ymat, rmat = rmat, cmat = cmat, wmat = wmat, verbose = FALSE)
h$itel
h$loss
h$x
```

## Nested, inmax = 1

```{r}
h <- nested(ymat, rmat = rmat, cmat = cmat, wmat = wmat, inmax = 1, verbose = FALSE)
h$itel
h$loss
h$x
```

## Nested, inmax = 5

```{r}
h <- nested(ymat, rmat = rmat, cmat = cmat, wmat = wmat, inmax = 5, verbose = FALSE)
h$itel
h$loss
h$x
```

## Nested, inmax = 10

```{r}
h <- nested(ymat, rmat = rmat, cmat = cmat, wmat = wmat, inmax = 10, verbose = FALSE)
h$itel
h$loss
h$x
```

## Nested, inmax = 100

```{r}
h <- nested(ymat, rmat = rmat, cmat = cmat, wmat = wmat, inmax = 100, verbose = FALSE)
h$itel
h$loss
h$x
```

\sectionbreak

# References