% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother


% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}



\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}



 


\usepackage{tcolorbox}
\usepackage{amssymb}
\usepackage{yfonts}
\usepackage{bm}


\newtcolorbox{greybox}{
  colback=white,
  colframe=blue,
  coltext=black,
  boxsep=5pt,
  arc=4pt}
  
\newcommand{\sectionbreak}{\clearpage}

 
\newcommand{\ds}[4]{\sum_{{#1}=1}^{#3}\sum_{{#2}=1}^{#4}}
\newcommand{\us}[3]{\mathop{\sum\sum}_{1\leq{#2}<{#1}\leq{#3}}}

\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\ul}[1]{\underline{#1}}

\newcommand{\amin}[1]{\mathop{\text{argmin}}_{#1}}
\newcommand{\amax}[1]{\mathop{\text{argmax}}_{#1}}

\newcommand{\ci}{\perp\!\!\!\perp}

\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\mb}[1]{\mathbb{#1}}
\newcommand{\mf}[1]{\mathfrak{#1}}

\newcommand{\eps}{\epsilon}
\newcommand{\lbd}{\lambda}
\newcommand{\alp}{\alpha}
\newcommand{\df}{=:}
\newcommand{\am}[1]{\mathop{\text{argmin}}_{#1}}
\newcommand{\ls}[2]{\mathop{\sum\sum}_{#1}^{#2}}
\newcommand{\ijs}{\mathop{\sum\sum}_{1\leq i<j\leq n}}
\newcommand{\jis}{\mathop{\sum\sum}_{1\leq j<i\leq n}}
\newcommand{\sij}{\sum_{i=1}^n\sum_{j=1}^n}
	
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Matrix Approximation with Kronecker and Precision Weighting},
  pdfauthor={Jan de Leeuw; Jan Graffelman},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}


\title{Matrix Approximation with Kronecker and Precision Weighting}
\author{Jan de Leeuw \and Jan Graffelman}
\date{April 23, 2026}
\begin{document}
\maketitle
\begin{abstract}
Majorization (or MM) algorithms for matrix approximation, using
minimization of triple-weighted least squares loss functions, are
developed. Loss functions combine matrix variate weights (or Kronecker
weights) with elementwise (or precision) weights. Implementations in R
are provided.
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\setcounter{tocdepth}{3}
\tableofcontents
}

\sectionbreak

\textbf{Note:} This is a working manuscript which will be
expanded/updated frequently. All suggestions for improvement are
welcome. All Rmd, tex, html, pdf, R, and C files are in the public
domain. Attribution will be appreciated, but is not required. The files
can be found at \url{https://github.com/deleeuw/triple}

\sectionbreak

\section{Introduction}\label{introduction}

The problem addressed in this paper is to approximate a given
\(n\times m\) matrix \(Y\) by a \(n\times m\) matrix \(X\) which must be
in a given subset \(\mathbb{X}\) of \(\mathbb{R}^{n\times m}\). For
example, \(\mathbb{X}\) can be the \(n\times m\) matrices of rank less
than or equal to \(p\), or a set of special matrices, such as the Hankel
or Cauchy matrices.

We measure loss using a loss
function\footnote{The symbol $:=$ is used for definitions. Symbol $\circ$ is used for Hadamard (elementwise) multiplication.}
proposed in Graffelman (\citeproc{ref-graffelman_26a}{2026b}).
\begin{equation}
\sigma(X):=\text{tr}\ R(W\circ(Y-X))C(W\circ(Y-X))'.\label{eq-loss}
\end{equation} In definition \eqref{eq-loss} \(R\) and \(C\) are given
positive semi-definite matrices of orders \(n\) and \(m\), which define
the \emph{Kronecker weights}. The Kronecker weights can be, for example,
inverse covariance matrices for row and column entries, as in the matrix
variate normal distribution (Gupta and Nagar
(\citeproc{ref-gupta_nagar_00}{2000})). The \(n\times m\) matrix \(W\)
of \emph{precision weights} has positive entries, which can be, for
example, inverse standard errors of the corresponding entries of \(Y\).

Loss function \eqref{eq-loss} has many well-known special cases. If all
weights are equal to one then we recover correspondence analysis,
redundancy analysis, and various other forms of canonical analysis, as
well as forms of principal component analysis with Kronecker weights (De
Leeuw (\citeproc{ref-deleeuw_A_84b}{1984}), De Leeuw and Kukuyeva
(\citeproc{ref-deleeuw_kukuyeva_U_09}{2009}), Graffelman
(\citeproc{ref-graffelman_26a}{2026b}), Graffelman
(\citeproc{ref-graffelman_26b}{2026a})). If both \(R\) and \(C\) are
equal to the identity we have principal components analysis with
precision weights, as in Gabriel and Zamir
(\citeproc{ref-gabriel_zamir_79}{1979}). If \(Y\) is symmetric the
various approximations of the correlation matrix discussed in
{Graffelman and de Leeuw} (\citeproc{ref-graffelman_deleeuw_A_23}{2023})
are special cases.

\sectionbreak

\section{Algorithms}\label{algorithms}

Our algorithms use majorization. See Lange
(\citeproc{ref-lange_16}{2016}) or De Leeuw
(\citeproc{ref-deleeuw_B_21c}{2022}) for comprehensive discussions of
majorization (or MM). In this paper majorization is used to reduce the
weighted least squares problem to an unweighted least squares problem,
which is generally much easier to solve. Similar applications of MM are
in Kiers (\citeproc{ref-kiers_97}{1997}) and in Groenen et al.
(\citeproc{ref-groenen_giaquinto_kiers_03}{2003}).

\subsection{Simple Majorization}\label{simple-majorization}

We first \emph{vectorize} the problem. This means replacing definition
\eqref{eq-loss} that uses matrices by an equivalent definition in terms
of vectors. Vectors are easier to work with. Define \(x:=\text{vec}(X)\)
and \(y:=\text{vec}(X)\). Loss becomes \begin{equation}
\sigma(x):=(y-x)'W_D(C\otimes R)W_D(y-x).\label{eq-lossv}
\end{equation} In \eqref{eq-lossv} matrix \(W_D\) is the diagonal matrix
with the elements of \(\text{vec}(W)\) on the diagonal.

Suppose \(\lambda\) is any number larger than or equal to the largest
eigenvalue \(\lambda_{\text{max}}(H)\) of the matrix \begin{equation}
H:=W_D(C\otimes R)W_D.\label{eq-hdef}
\end{equation} Suppose \(\tilde x\) is another vector of the same length
as \(x\). Use \(x=\tilde x+(x-\tilde x)\) in \eqref{eq-lossv}. For all
\(x\) and \(\tilde x\) we
have\footnote{We use $\|.\|^2$ for the (unweighted)
sum of squares.}, \begin{align}
\sigma(x)&=((y-\tilde x)-(x-\tilde x))'H((y-\tilde x)-(x-\tilde x))\notag\\
&=\sigma(\tilde x)-2(x-\tilde x)'H(y-\tilde x)+(x-\tilde x)'H(x-\tilde x)\notag\\
&\leq\sigma(\tilde x)-2(x-\tilde x)'H(y-\tilde x)+\lambda\|x-\tilde x\|^2\notag\\
&=\sigma(\tilde x)+\lambda\|x-\tilde y\|^2-\lambda^{-1}\|H(y-\tilde x)\|^2,
\end{align} with \begin{equation}
\tilde y:=\tilde x+\lambda^{-1}H(y-\tilde x).\label{eq-ytrip}
\end{equation}

A majorization (MM) step minimizes \(\|x-\tilde y\|^2\) over the \(x\)
satisfying the constraints. If we now \emph{matricize} (the inverse of
vectorize) definition \eqref{eq-ytrip} it follows that we have to
minimize \(\|X-\tilde Y\|^2\) over \(\mathbb{X}\), with \begin{equation}
\tilde Y:=\tilde X+\lambda^{-1}W\circ\{R(W\circ(Y-\tilde X))C\}.
\end{equation} Note that this expression for \(\tilde Y\) does not use
the (possibly very large) Kronecker product \(H\). If computing the
largest eigenvalue \(\lambda_{\text{max}}(H)\) (which only has to be
done once before the iterations start) is a problem then we can use an
upper bound. \begin{multline}
\max_z\frac{z'W_D(C\otimes R)W_Dz}{z'z}=\max_z\frac{z'W_D(C\otimes R)W_Dz}{z'W_D^2z}\frac{z'W_D^2z}{z'z}\\
\leq\max_z\frac{z'W_D(C\otimes R)W_Dz}{z'W_D^2z}\max_z\frac{z'W_D^2z}{z'z}
=\lambda_{\text{max}}(C)\lambda_{\text{max}}(R)\max(W\circ W)
\end{multline}

\subsection{Nested Majorization}\label{nested-majorization}

There is a more circumspect way to avoid computing the largest
eigenvalue of \(H\). We employ two nested majorizations. In the first
majorization we get rid of the Kronecker weighting and reduce each MM
iteration to a weighted least squares problem that is separable,
i.e.~has a diagonal weight matrix. The second majorization is used to
make a further reduction of this separable weighted least squares
problem to a sequence of unweighted least squares problems. The
majorizations are nested, in the sense that we need one or more
iterations of the second majorization process within each iteration of
the first majorization process.

\subsubsection{First Majorization}\label{first-majorization}

Start with the same vectorization we used before in our simple
majorization But instead of majorizing by using
\(z'Hz\leq\lambda_{\text{max}}(H)z'z\) we use \begin{equation}
z'Hz\leq\lambda_{\text{max}}(C\otimes R)z'W_D^2z=\lambda_{\text{max}}(C)\lambda_{\text{max}}(R)z'W_D^2z.
\end{equation} Let
\(\lambda:=\lambda_{\text{max}}(C)\lambda_{\text{max}}(R)\). We now have
\begin{align}
\sigma(x)&=((y-\tilde x)-(x-\tilde x))'H((y-\tilde x)-(x-\tilde x))\notag\\
&=\sigma(\tilde x)-2(x-\tilde x)'H(y-\tilde x)+(x-\tilde x)'H(x-\tilde x)\notag\\
&\leq\sigma(\tilde x)-2(x-\tilde x)'H(y-\tilde x)+\lambda(x-\tilde x)'W_D^2(x-\tilde x)\notag\\
&=\sigma(\tilde x)+\lambda(x-\tilde y)'W_D^2(x-\tilde y)+\lambda^{-1}(y-\tilde x)'HW_D^{-2}H(y-\tilde x)
\end{align} The target for the projection becomes \begin{equation}
\tilde y:=\tilde x+\lambda^{-1}W_D^{-2}H(y-\tilde x),
\end{equation} which can be matricized to \begin{equation}
\tilde Y:=\tilde X+\lambda^{-1}W^{-1}\circ\{R(W\circ(Y-\tilde X))C\}.\label{eq-ynest}
\end{equation} Note that in definition \eqref{eq-ynest} \(W^{-1}\) is
the elementwise inverse of \(W\).

\subsubsection{Second majorization}\label{second-majorization}

In the ``inner'' majorization we need to minimize \begin{equation}
S_W(X):-\sum_{i=1}^n\sum_{j=1}^nw_{ij}^2(\tilde y_{ij}-x_{ij})^2\label{eq-inloss}
\end{equation} over \(X\in\mathbb{X}\). Note that in the inner
iterations the ``target'' \(\tilde Y\) is the same in each iteration,
because it only changes in the outer iterations. Minimizing (or
decreasing) \eqref{eq-inloss} can be done with the wAddPCA() function
from the Correlplot package ({Graffelman and de Leeuw}
(\citeproc{ref-graffelman_deleeuw_M_26}{2026})), but we for ease of
reference we give the majorization for the inner iterations.

The same quadratic majorization reasoning as in the previous sections
leads to the unweighted problem of minimizing \begin{equation}
S(X):=\|\tilde{\tilde Y}-X\|^2,
\end{equation} with \begin{equation}
\tilde{\tilde y}_{ij}:=\tilde x_{ij}+\frac{w_{ij}^2}{w^2_{\text{max}}}(\tilde y_{ij}-\tilde x_{ij}).
\end{equation} Inner iterations project \(\tilde{\tilde Y}\) on
\(\mathbb{X}\) using unweighted least squares.

\sectionbreak

\section{Software}\label{software}

The github repository at \url{https://github.com/deleeuw/triple} has the
file triple.R, which constains the function triple(), for simple
majorization, and the file nested.R with the R program nested(), for
nested majorization. The file auxiliary.R has routines for different
unweighted least squares projections in matrix space.

\subsection{Arguments}\label{arguments}

Arguments are

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{args}\NormalTok{(triple)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
function (ymat, wmat = array(1, dim(ymat)), rmat = diag(nrow(ymat)), 
    cmat = diag(ncol(ymat)), func = eckart_young, p = 2, xini = NULL, 
    itmax = 1e+05, eps = 1e-06, verbose = TRUE) 
NULL
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{args}\NormalTok{(nested)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
function (ymat, wmat = array(1, dim(ymat)), rmat = diag(nrow(ymat)), 
    cmat = diag(ncol(ymat)), func = eckart_young, p = 2, xini = NULL, 
    itmax = 1e+05, inmax = 10, eps = 1e-06, ips = 1e-06, verbose = TRUE) 
NULL
\end{verbatim}

The arguments are mostly self-explanotory. An exception is perhaps the
func parameter. which is a separate R routine that performs the
unweighted least squares projection on \(\mathbb{X}\). Parameter \(p\)
is a vector or list of additional parameters for func, if needed. The
default func uses the truncated SVD for low-rank matrix approximation
with \(p=2\). If xini is NULL (the default) then the initial estimate
for both triple() and nested() is the unweighted projection of the data
matrix, without taking any of the weight matrices into account. If xini
is non-NULL it is used as the initial estimate. If verbose is TRUE then
information for each iteration is written to stdout. Otherwise the
programs are silent until they are done. Note that nested() has stopping
parameters itmax and eps for the outer iterations and inmax and ips for
the inner iterations.

\subsection{Values}\label{values}

Both triple() and nested return a list with three elements.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{list}\NormalTok{(}\AttributeTok{x =}\NormalTok{ xnew, }\AttributeTok{loss =}\NormalTok{ snew, }\AttributeTok{itel =}\NormalTok{ itel)}
\end{Highlighting}
\end{Shaded}

If the verbose parameter is true then triple() writes itel, sold, snew,
and epsi to stdout. Here epsi is the sup-norm of the difference between
the current and the previous solution of \(X\). For nested() the
iteration output also includes ntel (the number of inner iterations in
the current outer iteration) and ipsi (the sup-norm of the difference
between xold and xnew for the last inner iteration).

\subsection{Projection Routines}\label{projection-routines}

The routines in auxiliary.R define possible values for the func
parameter of triple() and nested(). Users can add their own functions as
needed. Currently we have

\begin{itemize}
\tightlist
\item
  eckart\_young() low-rank approximation.
\item
  column\_adjust() subtract column means and then eckart\_young.
\item
  row\_adjust() subtract row means and then eckart\_young
\item
  double\_center() and then eckart\_young.
\item
  hankel() finds the best fitting Hankel matrix (square matrices only).
\item
  toeplitz() finds the best fitting Toeplitz matrix (square matrices
  only).
\item
  symmetric() finds the best fitting symmetric matrix (square matrices
  only).
\item
  anti\_symmetric() finds the best fitting anti-symmetric matrix (square
  matrices only).
\item
  additive() finds the best fitting additive matrix (i.e.~with
  \(x_{ij}=a_i+b_j\)).
\end{itemize}

\sectionbreak

\section{Examples}\label{examples}

\subsection{Artificial}\label{artificial}

We analyze a small artifical example, generated from random data.

\subsubsection{Data}\label{data}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{ymat}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
           [,1]       [,2]       [,3]
[1,]  0.5855288  0.6058875 -0.2841597
[2,]  0.7094660 -1.8179560 -0.9193220
[3,] -0.1093033  0.6300986 -0.1162478
[4,] -0.4534972 -0.2761841  1.8173120
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rmat}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
          [,1]       [,2]       [,3]       [,4]
[1,] 0.6606540  0.6316750  0.2060831  0.4388240
[2,] 0.6316750  1.5141559 -0.7485539  0.5940988
[3,] 0.2060831 -0.7485539  2.0102613 -0.4442019
[4,] 0.4388240  0.5940988 -0.4442019  1.3318363
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cmat}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
           [,1]       [,2]       [,3]
[1,]  1.3744581  0.7610956 -0.2692483
[2,]  0.7610956  1.2428719 -0.1845269
[3,] -0.2692483 -0.1845269  1.3504256
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{wmat}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
           [,1]       [,2]      [,3]
[1,] 0.09264056 0.45164108 0.9291061
[2,] 6.13607849 0.09483528 0.7311661
[3,] 0.94326960 0.28785770 3.5605688
[4,] 3.48605937 0.68041062 0.1535224
\end{verbatim}

\subsubsection{Triple}\label{triple}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h }\OtherTok{\textless{}{-}} \FunctionTok{triple}\NormalTok{(ymat, }\AttributeTok{rmat =}\NormalTok{ rmat, }\AttributeTok{cmat =}\NormalTok{ cmat, }\AttributeTok{wmat =}\NormalTok{ wmat, }\AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{)}
\NormalTok{h}\SpecialCharTok{$}\NormalTok{itel}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 28251
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h}\SpecialCharTok{$}\NormalTok{loss}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 0.0006645135
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h}\SpecialCharTok{$}\NormalTok{x}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
            [,1]       [,2]       [,3]
[1,] -0.07238101  0.7118148 -0.2867966
[2,]  0.71448655 -2.1034735 -0.9179862
[3,] -0.08977028  0.5685101 -0.1163371
[4,] -0.45002486 -0.2906866  1.8085130
\end{verbatim}

\subsubsection{Nested, inmax = 1}\label{nested-inmax-1}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h }\OtherTok{\textless{}{-}} \FunctionTok{nested}\NormalTok{(ymat, }\AttributeTok{rmat =}\NormalTok{ rmat, }\AttributeTok{cmat =}\NormalTok{ cmat, }\AttributeTok{wmat =}\NormalTok{ wmat, }\AttributeTok{inmax =} \DecValTok{1}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{)}
\NormalTok{h}\SpecialCharTok{$}\NormalTok{itel}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 56285
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h}\SpecialCharTok{$}\NormalTok{loss}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 0.0006705183
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h}\SpecialCharTok{$}\NormalTok{x}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
            [,1]       [,2]       [,3]
[1,] -0.07268466  0.7098735 -0.2856829
[2,]  0.71435241 -2.0877511 -0.9185953
[3,] -0.09053565  0.5701966 -0.1165036
[4,] -0.45001529 -0.2910992  1.8029899
\end{verbatim}

\subsubsection{Nested, inmax = 5}\label{nested-inmax-5}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h }\OtherTok{\textless{}{-}} \FunctionTok{nested}\NormalTok{(ymat, }\AttributeTok{rmat =}\NormalTok{ rmat, }\AttributeTok{cmat =}\NormalTok{ cmat, }\AttributeTok{wmat =}\NormalTok{ wmat, }\AttributeTok{inmax =} \DecValTok{5}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{)}
\NormalTok{h}\SpecialCharTok{$}\NormalTok{itel}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 11292
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h}\SpecialCharTok{$}\NormalTok{loss}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 0.0006705229
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h}\SpecialCharTok{$}\NormalTok{x}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
            [,1]       [,2]       [,3]
[1,] -0.07268506  0.7098702 -0.2856778
[2,]  0.71435237 -2.0877448 -0.9185993
[3,] -0.09053604  0.5701999 -0.1165041
[4,] -0.45001522 -0.2910987  1.8029850
\end{verbatim}

\subsubsection{Nested, inmax = 10}\label{nested-inmax-10}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h }\OtherTok{\textless{}{-}} \FunctionTok{nested}\NormalTok{(ymat, }\AttributeTok{rmat =}\NormalTok{ rmat, }\AttributeTok{cmat =}\NormalTok{ cmat, }\AttributeTok{wmat =}\NormalTok{ wmat, }\AttributeTok{inmax =} \DecValTok{10}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{)}
\NormalTok{h}\SpecialCharTok{$}\NormalTok{itel}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 5674
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h}\SpecialCharTok{$}\NormalTok{loss}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 0.0006705318
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h}\SpecialCharTok{$}\NormalTok{x}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
            [,1]       [,2]       [,3]
[1,] -0.07268565  0.7098653 -0.2856708
[2,]  0.71435231 -2.0877316 -0.9186048
[3,] -0.09053672  0.5702046 -0.1165048
[4,] -0.45001520 -0.2910982  1.8029762
\end{verbatim}

\subsubsection{Nested, inmax = 100}\label{nested-inmax-100}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h }\OtherTok{\textless{}{-}} \FunctionTok{nested}\NormalTok{(ymat, }\AttributeTok{rmat =}\NormalTok{ rmat, }\AttributeTok{cmat =}\NormalTok{ cmat, }\AttributeTok{wmat =}\NormalTok{ wmat, }\AttributeTok{inmax =} \DecValTok{100}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{)}
\NormalTok{h}\SpecialCharTok{$}\NormalTok{itel}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 634
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h}\SpecialCharTok{$}\NormalTok{loss}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 0.0006707031
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h}\SpecialCharTok{$}\NormalTok{x}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
            [,1]       [,2]       [,3]
[1,] -0.07269601  0.7097798 -0.2855324
[2,]  0.71435159 -2.0874979 -0.9187098
[3,] -0.09054544  0.5702865 -0.1165185
[4,] -0.45001542 -0.2910846  1.8027803
\end{verbatim}

\subsubsection{Conclusion}\label{conclusion}

The artificial example is a minimal form of quality control. It shows us
that triple() and the various nested() versions converge (at least stop
before reaching the 100,000 iteration limit), and all converge to the
same solution (when started at the same initial point and when using the
same stop criteria).

The example also shows us that convergence can be very slow and close to
sublinear. Function triple() uses about 30,000 iterations, and nested()
with a single inner iteration uses about 60,000 iterations. That means,
in this case, 30,000 or 60,000 Eckart-Young decompositions. Ouch !

\subsection{Correspondence Analysis}\label{correspondence-analysis}

In correspondence analysis and multiple correspondence analysis the
Kronecker weight are the diagonal matrices with the reciprocals of the
marginals of the table.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(anacor)}
\FunctionTok{data}\NormalTok{(galton)}
\NormalTok{ymat }\OtherTok{\textless{}{-}} \FunctionTok{as.matrix}\NormalTok{(galton) }\SpecialCharTok{/} \FunctionTok{sum}\NormalTok{(galton)}
\NormalTok{rmat }\OtherTok{\textless{}{-}} \FunctionTok{diag}\NormalTok{(}\DecValTok{1} \SpecialCharTok{/} \FunctionTok{rowSums}\NormalTok{(ymat))}
\NormalTok{cmat }\OtherTok{\textless{}{-}} \FunctionTok{diag}\NormalTok{(}\DecValTok{1} \SpecialCharTok{/} \FunctionTok{colSums}\NormalTok{(ymat))}
\NormalTok{wmat }\OtherTok{\textless{}{-}} \FunctionTok{array}\NormalTok{(}\DecValTok{1}\NormalTok{, }\FunctionTok{dim}\NormalTok{(ymat))}
\NormalTok{h }\OtherTok{\textless{}{-}} \FunctionTok{triple}\NormalTok{(ymat, wmat, rmat, cmat, }\AttributeTok{p =} \DecValTok{2}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{)}
\FunctionTok{print}\NormalTok{(h}\SpecialCharTok{$}\NormalTok{x)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
               [,1]          [,2]          [,3]          [,4]          [,5]
 [1,] -6.179741e-05 -0.0001365276 -0.0004971256 -0.0004069247 -0.0002991340
 [2,] -1.267911e-04 -0.0003165737 -0.0010949673 -0.0005976866 -0.0003944022
 [3,]  1.870542e-05 -0.0001076505 -0.0001560155  0.0010924909  0.0009868318
 [4,]  1.859950e-04  0.0001396979  0.0009382487  0.0029894275  0.0025320462
 [5,]  8.209118e-04  0.0010146654  0.0049600739  0.0106039990  0.0087804464
 [6,]  1.412997e-03  0.0020667578  0.0091964227  0.0161683744  0.0131865786
 [7,]  1.545993e-03  0.0023645165  0.0102743917  0.0170185180  0.0138066732
 [8,]  6.474517e-04  0.0010272181  0.0043789158  0.0068866742  0.0055596998
 [9,]  5.419081e-04  0.0008594413  0.0036644212  0.0057661723  0.0046553533
[10,]  2.289664e-04  0.0003795025  0.0015819700  0.0023297882  0.0018684725
[11,]  1.348105e-04  0.0002218764  0.0009282071  0.0013819224  0.0011095416
               [,6]          [,7]         [,8]         [,9]       [,10]
 [1,] -0.0008281501 -0.0005531613 0.0002258172 0.0007554199 0.001219873
 [2,] -0.0012439258 -0.0002786423 0.0017382758 0.0036668940 0.004305928
 [3,]  0.0021108057  0.0036665292 0.0051415699 0.0084220450 0.006998779
 [4,]  0.0058789756  0.0080351181 0.0088052164 0.0134753193 0.009742258
 [5,]  0.0209769478  0.0261137368 0.0249410025 0.0363586592 0.023309903
 [6,]  0.0321078183  0.0374261765 0.0317297437 0.0439855989 0.024282827
 [7,]  0.0338409561  0.0385242107 0.0311061307 0.0421313255 0.021462880
 [8,]  0.0137107153  0.0152652649 0.0117340164 0.0154973427 0.007159874
 [9,]  0.0114797497  0.0127844731 0.0098326111 0.0129899932 0.006008840
[10,]  0.0046459817  0.0050171305 0.0035818936 0.0045378058 0.001729104
[11,]  0.0027550149  0.0029907740 0.0021637276 0.0027627301 0.001095538
             [,11]         [,12]         [,13]         [,14]
 [1,] 0.0008993253  0.0019137376  0.0012009759  0.0008292192
 [2,] 0.0030951851  0.0057035669  0.0034995251  0.0024367387
 [3,] 0.0048357843  0.0066825988  0.0038677006  0.0027541423
 [4,] 0.0065925695  0.0074606366  0.0040884674  0.0029752270
 [5,] 0.0154475706  0.0135233556  0.0067383251  0.0051012328
 [6,] 0.0156082473  0.0076659056  0.0025022822  0.0023202111
 [7,] 0.0135378584  0.0033554850 -0.0001941327  0.0004562490
 [8,] 0.0044018325 -0.0003970400 -0.0011314330 -0.0005547599
 [9,] 0.0036954516 -0.0003164245 -0.0009377338 -0.0004577496
[10,] 0.0010000296 -0.0009317729 -0.0008612207 -0.0005236718
[11,] 0.0006425096 -0.0004722451 -0.0004625745 -0.0002767260
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{print}\NormalTok{(h}\SpecialCharTok{$}\NormalTok{itel)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 1417
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{print}\NormalTok{(h}\SpecialCharTok{$}\NormalTok{loss)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 0.2824752
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{svd}\NormalTok{(h}\SpecialCharTok{$}\NormalTok{x)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
$d
 [1] 1.412689e-01 2.300972e-02 1.201181e-17 8.849489e-18 5.180446e-18
 [6] 2.947762e-18 1.737753e-18 1.083998e-18 6.271810e-19 2.590316e-19
[11] 1.087156e-19

$u
             [,1]        [,2]        [,3]        [,4]        [,5]          [,6]
 [1,] -0.00384317 -0.13696824  0.11596875 -0.09529819  0.09238202  0.0246153407
 [2,] -0.03109658 -0.38739982  0.01913720  0.74839884 -0.08371732 -0.5118014176
 [3,] -0.09368657 -0.39324181 -0.15861113  0.08048632 -0.17198732  0.5946444000
 [4,] -0.16111093 -0.37914900 -0.07163202  0.17747645 -0.25016025  0.4195960313
 [5,] -0.45762607 -0.51181852  0.13396641 -0.56914543 -0.13044275 -0.3869083713
 [6,] -0.58378676  0.05352947  0.28775390  0.21021645  0.69269320  0.1922430121
 [7,] -0.57301021  0.35978444 -0.67742854  0.06872039 -0.19784397 -0.1147834139
 [8,] -0.21643256  0.25856214  0.46632706  0.07374015 -0.47675988  0.0534998732
 [9,] -0.18135863  0.21546159  0.38905173  0.11789986 -0.32240539  0.0424782341
[10,] -0.06620345  0.13881008  0.14885411  0.02362283 -0.08585612  0.0573822184
[11,] -0.03997658  0.07715727  0.07251154  0.01027117 -0.13303857  0.0007420257
              [,7]         [,8]        [,9]       [,10]         [,11]
 [1,]  0.165326864 -0.702877302  0.34889802  0.43184492  3.463593e-01
 [2,]  0.073489607  0.037667629 -0.02823437  0.08373251  7.007253e-02
 [3,]  0.522844398  0.231844065 -0.21180833  0.17572111  1.375062e-01
 [4,] -0.611601667 -0.212733021  0.24932607 -0.21954601 -1.591961e-01
 [5,]  0.009206413  0.134922823 -0.01488735 -0.03197625 -2.236564e-02
 [6,] -0.022743601  0.021716876 -0.08574031 -0.06490618 -1.455649e-05
 [7,]  0.048932975 -0.106924694  0.05691808  0.09030234  3.230435e-02
 [8,] -0.142468438 -0.208429352 -0.55424798  0.24885325 -1.957018e-02
 [9,]  0.441264932  0.045396190  0.55359854 -0.27007869 -2.710806e-01
[10,] -0.318942519  0.574365524  0.37862289  0.50375680  3.379689e-01
[11,] -0.010518992 -0.005440064 -0.05115238 -0.57042907  8.007742e-01

$v
             [,1]         [,2]         [,3]        [,4]        [,5]       [,6]
 [1,] -0.01679722  0.022501887 -0.192488455 -0.31528860  0.41761468 -0.5549279
 [2,] -0.02435081  0.047514940 -0.956537002  0.13343457  0.02912334  0.1310748
 [3,] -0.10887519  0.176493228 -0.078890832  0.07067146 -0.07228970  0.0344136
 [4,] -0.19362291  0.162471889  0.097532586  0.17718746 -0.22438911 -0.1313906
 [5,] -0.15806946  0.122145262 -0.024141273  0.01097973 -0.10125169 -0.4099963
 [6,] -0.38440945  0.328992759 -0.019419980 -0.04547187 -0.52695968 -0.2006090
 [7,] -0.45003143  0.253043649  0.130863832  0.35723446  0.63125225 -0.1209780
 [8,] -0.38481864 -0.005359606 -0.062993962  0.17145611 -0.12922111  0.1305799
 [9,] -0.53554876 -0.147436059 -0.013803126 -0.73153896  0.08230582  0.2961653
[10,] -0.29945146 -0.335478662  0.058753802  0.35605948  0.10526484  0.3516249
[11,] -0.19302275 -0.252103427 -0.067099466  0.11323831 -0.20337977 -0.2767715
[12,] -0.10175993 -0.589697174 -0.044268234  0.10651043 -0.05453315 -0.2306840
[13,] -0.03594016 -0.374877194  0.007913174  0.01567765 -0.04357418 -0.2282733
[14,] -0.03198095 -0.257601199 -0.005301277  0.02925356 -0.04434856 -0.1520485
             [,7]        [,8]        [,9]       [,10]       [,11]
 [1,]  0.03094868 -0.49411355 -0.20774043  0.12057215  0.25015522
 [2,] -0.07478703  0.15772191 -0.07526963 -0.05841715 -0.06549740
 [3,]  0.06001484 -0.32354554  0.79321414 -0.20247185  0.16307286
 [4,] -0.01837551 -0.08270196 -0.41477490 -0.16690691 -0.34194281
 [5,]  0.25237966  0.11752584 -0.01804408 -0.69703620 -0.02018361
 [6,] -0.31682781  0.13782210 -0.06520958  0.24001163  0.45996605
 [7,] -0.12519225  0.33889907  0.13144910  0.13106197 -0.09575370
 [8,]  0.77098539 -0.19849858 -0.07844820  0.31245107 -0.03631787
 [9,] -0.04961193  0.10254914  0.03232162 -0.11410184 -0.15838837
[10,] -0.26124763 -0.40391903 -0.24518078 -0.32816987  0.34034338
[11,] -0.33926397 -0.34076325  0.20670733  0.20198162 -0.59707770
[12,]  0.03695614  0.26908588  0.09982688  0.19728519  0.25994295
[13,]  0.03135665  0.23163702  0.07563076 -0.23101319 -0.01218898
[14,]  0.16486969  0.12862087  0.04867236  0.03360441  0.04483004
\end{verbatim}

\subsection{Canonical Analysis}\label{canonical-analysis}

\subsection{Triple Weighting}\label{triple-weighting}

\sectionbreak

\section{Discussion}\label{discussion}

The complications in minimizing loss function \eqref{eq-loss} arise from
the combination of the \((R,C)\) and the \(W\) weights. If either one of
these two sets of weights is removed (or set equal to the identity)
minimization becomes much easier. One of the two majorizations in the
nested process disappears. There is no need for the largest eigenvalue
of the large matrix \(H\) from \eqref{eq-hdef}. The same thing is true
if we replace loss function \eqref{eq-loss} by \begin{equation}
\sigma(X):=\text{tr}\ R(W\circ(Y)-X)C(W\circ(Y)-X)'.\label{eq-lossno}
\end{equation} In that case we also only need the first majorization
applied to \(W\circ(Y)\).

The distinction between simple an nested majorization is relative. In
simple majorization we may need to solve an eigenvalue problem in each
iteration, and the eigenvalue computation itself defines an inner
iteration. In that sense simple majorization is also nested, we just
call it simple because eigenvalue computations are implemented as black
boxes that had the iteration details. We could implement our projections
as explicit iterative processes with update rules and stop criteria. In
that case our nested majorization will have outer, inner, and innermost
iterations and three sets of stop criteria.

In many previous majorization methods the stop criterion is defined
using a decrease in the loss function from one iteration to the next of
less than epsilon. In triple() and nested() we prefer to use the
sup-norm of the change in the solution from one iteration to the next.
This is a stricter criterion, because solutions generally converge
slower than loss functions. It is also more satisfactory, because change
in the solution measures basically how well the conditions for
stationarity (vanishing of the gradient or the Karush-Kuhn-Tucker
conditions in the constrained case) are satisfied. By computing the
change in the projections we also circumvent identifiabilty problems. If
the Eckart-Young projection, for example, is parametrized as \(X=AB'\),
then we should not look at the change in \(A\) and \(B\) from one
iteration to the next, but in the change of \(AB'\). In MDS we should
not look at the change in the configuration, but in the change of the
distances.

For large examples that would be impractical. It suggests that
acceleration techniques are needed. As currently implemented all
projections, such as Eckart-Young, have a cold start, i.e.~do not use
projections in previous iterations. This is clearly wasteful.

\sectionbreak

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\protect\phantomsection\label{refs}
\begin{CSLReferences}{1}{1}
\bibitem[\citeproctext]{ref-deleeuw_A_84b}
De Leeuw, Jan. 1984. {``Fixed-Rank Approximation with Singular Weight
Matrices.''} \emph{Computational Statistics Quarterly} 1: 3--12.
\url{https://jansweb.netlify.app/publication/deleeuw-a-84-b/deleeuw-a-84-b.pdf}.

\bibitem[\citeproctext]{ref-deleeuw_B_21c}
De Leeuw, Jan. 2022. \emph{Block Relaxation Methods in Statistics}.
Bookdown.
\url{https://jansweb.netlify.app/publication/deleeuw-b-21-c/deleeuw-b-21-c.pdf}.

\bibitem[\citeproctext]{ref-deleeuw_kukuyeva_U_09}
De Leeuw, Jan, and Irina Kukuyeva. 2009. {``{Fixed-Rank Approximation
with Constraints}.''}
\url{https://jansweb.netlify.app/publication/deleeuw-kukuyeva-u-09/deleeuw-kukuyeva-u-09.pdf}.

\bibitem[\citeproctext]{ref-gabriel_zamir_79}
Gabriel, K. Ruben, and S. Zamir. 1979. {``{Lower Rank Approximation of
Matrices by Least Squares with Any Choice of Weights}.''}
\emph{Technometrics} 21 (4): 489--98.

\bibitem[\citeproctext]{ref-graffelman_26b}
Graffelman, Jan. 2026a. {``On the Approximation of the Between-Set
Correlation Matrix by Canonical Correlation Analysis.''} Unpublished
manuscript.

\bibitem[\citeproctext]{ref-graffelman_26a}
Graffelman, Jan. 2026b. {``Precision-Weighted Compositional Data
Analysis by Weighted Alternating Least Squares and Majorization
Algorithms.''} Unpublished manuscript.

\bibitem[\citeproctext]{ref-graffelman_deleeuw_A_23}
{Graffelman, Jan, and Jan de Leeuw}. 2023. {``Improved Approximation and
Visualization of the Correlation Matrix.''} \emph{American Statistician}
77 (4): 432--42.

\bibitem[\citeproctext]{ref-graffelman_deleeuw_M_26}
{Graffelman, Jan, and Jan de Leeuw}. 2026. \emph{{Correlplot: A
Collection of Functions for Graphing Correlation Matrices}}.
\url{https://CRAN.R-project.org/package=Correlplot}.

\bibitem[\citeproctext]{ref-groenen_giaquinto_kiers_03}
Groenen, Patrick J. F., Patrizia Giaquinto, and Henk A. L Kiers. 2003.
\emph{{Weighted Majorization Algorithms for Weighted Least Squares
Decomposition Models}}. Econometric Institute Report EI 2003-09.
Econometric Institute, Erasmus University Rotterdam.
\url{https://repub.eur.nl/pub/1700}.

\bibitem[\citeproctext]{ref-gupta_nagar_00}
Gupta, Arjan K., and D. K. Nagar. 2000. \emph{Matrix Variate
Distributions}. {Chapman \& Hall}.

\bibitem[\citeproctext]{ref-kiers_97}
Kiers, Henk A. L. 1997. {``{Weighted Least Squares Fitting Using
Iterative Ordinary Least Squares Algorithms}.''} \emph{Psychometrika}
62: 251--66.

\bibitem[\citeproctext]{ref-lange_16}
Lange, Kenneth. 2016. \emph{MM Optimization Algorithms}. SIAM.

\end{CSLReferences}




\end{document}
